An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-10-13 aistats2017.tex
Minor typo changes.

\begin{abstract}
  %The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
  %both left and right-hand margins. Use 10~point type, with a vertical
  %spacing of 11~points. The {\bf Abstract} heading must be centered,
  %bold, and in point size 12. Two line spaces precede the
  %Abstract. The Abstract must be limited to one paragraph.
  We
  present a novel Metropolis-Hastings method for large datasets that
  uses small expected-size minibatches of data. Previous work on
  reducing the cost of Metropolis-Hastings tests yield variable data
  consumed per sample, with only constant factor reductions versus
  using the full dataset for each sample.  Here we present a method
  that can be tuned to provide arbitrarily small batch sizes, by
  adjusting either proposal step size or temperature. Our test uses
  the noise-tolerant Barker acceptance test with a novel additive
  correction variable.  The resulting test has simlar cost to a normal
  SGD update. Our experiments demonstrate several order-of-magnitude
  speedup over previous work.
\end{abstract}

@@

It is worth discussing at this point the typical goals of M-H sampling on large
datasets. By the Bernstein-von Mises Theorem, the posterior distribution for a
Bayesian inference task has variance that scales inversely with $N$. Simply
sampling from it is one application, but an efficient
proposal~\citep{OptimalScaling01} has similar variance to the target and will
diffuse to it extremely slowly. For applications to neural networks or models
where the posterior is multimodal~\citep{choromanska2014loss}, samplers will
likely get trapped in one of the modes. A common solution is to anneal the
sampler, running first at high temperatures to flatten the likelihood landscape.
This in turn reduces the variance of the log acceptance probability and allows
our test to be applied. Theses samples can cover the search space densely with
small steps rather than taking a few sparse steps towards an optimum. In this
mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
Descent. The goal in SGD is to make gradual progress to a posterior mode with
each step, taking small steps so that the cumulative displacement has
progressively lower variance.

@@

In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$,
we can write
\begin{equation}\label{eq:relationship}
    \Delta^* = \Delta + X_{\rm norm}, \quad X_{\rm norm} \sim \mathcal{\bar{N}}(0, \sigma^2(\Delta^*)),
\end{equation}
where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is
approximately normal with variance $\sigma^2(\Delta^*)$.  But to perform the
test in Equation~(\ref{eq:the_exact_test}) we want $\Delta + X$ for a logistic
random variable $X$ (call it $X_{\rm log}$ from now on). In~\citet{TallData15} it
was proposed to use $\Delta^*$ in a Barker test, and tolerate the fixed
error between logistic and normal distributions.

@@

We ran MCMC with with our test against the test
from~\citet{icml2014c1_bardenet14} and a simplified version
of~\citet{cutting_mh_2014}. Both methods use repeated testing which
must be discounted to bound overall test
accuracy. ~\citet{icml2014c1_bardenet14} conservatively partitions the
test error, while~\citet{cutting_mh_2014} used a complex dynamic
programming heuristic to estimate the overall test error. We did not
implement this heuristic and instead set the individual test error
equal to the overall test error. Thus our results for
~\citet{cutting_mh_2014} should be treated as lower bounds. All
methods were initialized with minibatch size 100.  For ~\citet{cutting_mh_2014} we
increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For
~\citet{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma = 1.5$ and
use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
samples using a random walk proposer with covariance ${\rm diag}(0.15,
0.15)$, which means the M-H test is responsible for shaping the sample
distribution.

@@

We next test logistic regression for the binary classification of 1s
versus 7s on a subset of the MNIST8M dataset, a larger version of
MNIST~\citep{lecun-mnisthandwrittendigit-2010}. We randomly subsampled
450k training and 192k testing points.
%%% The pixels are scaled in $[0,1]$  - JFC Why ????
We impose a uniform prior on $\theta$ and again use a random walk
proposer, this time with covariance matrix $0.05I$. The temperature is
set at $T=1000$. We run the three tests for samples and again use 100
as the starting minibatch size. Method~\citet{cutting_mh_2014} used
per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
we used $\gamma=2.0$ (discounting factor). We tried to use the
symbolic bound for $C_{\theta,\theta'}$ for Logistic regression from
~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead
we used the empirical $C_{\theta,\theta'}$ from the entire dataset
without including the test for this test, so again our results
represent a lower bound for this method.

@@

\item Some reviewers mentioned that our deconvolution approach to
  determine the correction distribution was not unique and thus
  ill-defined.  We added detail description on how the correction
  distribution is computed. While the distribution is not unique, we
  regularize the inversion to obtain a concrete distribution with very
  low error (close to single-point machine precision).

@@

\item Our experiments now include new comparisons with a baseline
  from~\citep{icml2014c1_bardenet14} (in addition to~\citep{cutting_mh_2014},
  which we had earlier). In addition, each algorithm now runs MCMC sampling on
  the \emph{same} distribution.  Previously, we ran our distribution at a
  higher temperature but kept the algorithm from~\citep{cutting_mh_2014}
  running on the distribution at temperature $T=1$. In this set of
  experiments, we tuned hyperparameters of the other algorithms to yield
  best results. In the Gaussian mixture model scenario
  (Section~\ref{ssec:gaussians}), we provide likelihoods and test
  statistics to measure the accuracy of the samples.

>>>>

\begin{abstract}
  %The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
  %both left and right-hand margins. Use 10~point type, with a vertical
  %spacing of 11~points. The {\bf Abstract} heading must be centered,
  %bold, and in point size 12. Two line spaces precede the
  %Abstract. The Abstract must be limited to one paragraph.
  We
  present a novel Metropolis-Hastings method for large datasets that
  uses small expected-size minibatches of data. Previous work on
  reducing the cost of Metropolis-Hastings tests yield variable data
  consumed per sample, with only constant factor reductions versus
  using the full dataset for each sample.  Here we present a method
  that can be tuned to provide arbitrarily small batch sizes, by
  adjusting either proposal step size or temperature. Our test uses
  the noise-tolerant Barker acceptance test with a novel additive
  correction variable.  The resulting test has similar cost to a normal
  SGD update. Our experiments demonstrate order-of-magnitude
  speedups over previous work.
\end{abstract}

@@

It is worth discussing at this point the typical goals of M-H sampling on large
datasets. By the Bernstein-von Mises Theorem, the posterior distribution for a
Bayesian inference task has variance that scales inversely with $N$. Simply
sampling from it is one application, but an efficient
proposal~\citep{OptimalScaling01} has similar variance to the target and will
diffuse to it extremely slowly. For applications to neural networks or models
where the posterior is multimodal~\citep{choromanska2014loss}, samplers will
likely get trapped in one of the modes. A common solution is to anneal the
sampler, running first at high temperatures to flatten the likelihood landscape.
This in turn reduces the variance of the log acceptance probability and allows
our test to be applied. These samples can cover the search space densely with
small steps rather than taking a few sparse steps towards an optimum. In this
mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
Descent. The goal in SGD is to make gradual progress to a posterior mode with
each step, taking small steps so that the cumulative displacement has
progressively lower variance.

@@

In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$,
we can write
\begin{equation}\label{eq:relationship}
    \Delta^* = \Delta + X_{\rm norm}, \quad X_{\rm norm} \sim \mathcal{\bar{N}}(0, \sigma^2(\Delta^*)),
\end{equation}
where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is
approximately normal with variance $\sigma^2(\Delta^*)$.  But to perform the
test in Equation~(\ref{eq:the_exact_test}) we want $\Delta + X$ for a logistic
random variable $X$ (call it $X_{\rm log}$ from now on). In~\citet{TallData15} it
was proposed to use $\Delta^*$ in a Barker test, and tolerate the fixed
error between the logistic and normal distributions.

@@

We ran MCMC with with our test against the test
from~\citet{icml2014c1_bardenet14} and a simplified version
of~\citet{cutting_mh_2014}. Both methods use repeated testing which
must be discounted to bound overall test
accuracy. ~\citet{icml2014c1_bardenet14} conservatively partitions the
test error, while~\citet{cutting_mh_2014} used a complex dynamic
programming heuristic to estimate the overall test error. We did not
implement this heuristic and instead set the individual test error
equal to the overall test error. Thus our results for~\citet{cutting_mh_2014} should be treated as lower bounds. All
methods were initialized with minibatch size 100.  For~\citet{cutting_mh_2014} we
increment sizes by 100 and set the tolerance $\epsilon=0.005$.  For~\citet{icml2014c1_bardenet14}, we increase sizes geometrically with $\gamma = 1.5$ and
use parameters $p = 2, \delta = 0.01$.  All methods collect 5000
samples using a random walk proposer with covariance ${\rm diag}(0.15,
0.15)$, which means the M-H test is responsible for shaping the sample
distribution.

@@

We next test logistic regression for the binary classification of 1s
versus 7s on a subset of the MNIST8M dataset, a larger version of
MNIST~\citep{lecun-mnisthandwrittendigit-2010}. We randomly subsampled
450k training and 192k testing points.
%%% The pixels are scaled in $[0,1]$  - JFC Why ????
We impose a uniform prior on $\theta$ and again use a random walk
proposer, this time with covariance matrix $0.05I$. The temperature is
set at $T=1000$. We run the three tests for 3000 samples and again use 100
as the starting minibatch size. Method~\citet{cutting_mh_2014} used
per-test error of $\epsilon=0.05$.  For~\citet{icml2014c1_bardenet14},
we used $\gamma=2.0$ (discounting factor). We tried to use the
symbolic bound for $C_{\theta,\theta'}$ for Logistic regression
from~\citet{icml2014c1_bardenet14}, but found it to be too high. Instead we use
the empirical $C_{\theta,\theta'}$ from the entire dataset without including the
test for this test, so again our results represent a lower bound for this
method.

@@

\item Some reviewers mentioned that our deconvolution approach to
  determine the correction distribution was not unique and thus
  ill-defined.  We added detail on how the correction
  distribution is computed (see Section~\ref{sec:correction}). While the distribution is not unique, we
  regularize the inversion to obtain a concrete distribution with very
  low error (close to single-point machine precision).

@@

\item Our experiments now include new comparisons with a baseline
  from~\citep{icml2014c1_bardenet14} (in addition to~\citep{cutting_mh_2014},
  which we had earlier). Furthermore, each algorithm now runs MCMC sampling on
  the \emph{same} distribution.  Previously, we ran our distribution at a
  higher temperature but kept the algorithm from~\citep{cutting_mh_2014}
  running on the distribution at temperature $T=1$. In this set of
  experiments, we tuned hyperparameters of the other algorithms to yield
  best results. In the Gaussian mixture model scenario
  (Section~\ref{ssec:gaussians}), we provide likelihoods and test
  statistics to measure the accuracy of the samples.
